---
title: "Full TCGA dataframe with missing values"
output: html_notebook
---

## Purpose of the script

Data collection of the proteomic expression data and other features that
may be of interest of 28 cancer types.

**Input data sets:** Follow the [[Downloading the data from the GDC
website]] instructions for the following cancer studies

-   TCGA-GBM (Glioblastoma multiforme)

-   TCGA-LGG (Brain Lower Grade Glioma)

-   TCGA-HNSC (Head and Neck squamous cell carcinoma)

-   TCGA-THCA (Thyroid carcinoma)

-   TCGA-ESCA (Esophageal carcinoma)

-   TCGA-SARC (Sarcoma)

-   TCGA-SKCM (Skin Cutaneous Melanoma)

-   TCGA-LUAD (Lung adenocarcinoma)

-   TCGA-LUSC (Lung squamous cell carcinoma)

-   TCGA-MESO (Mesothelioma)

-   TCGA-THYM (Thymoma)

-   TCGA-BRCA (Breast invasive carcinoma)

-   TCGA-STAD (Stomach adenocarcinoma)

-   TCGA-LIHC (Liver hepatocellular carcinoma)

-   TCGA-PAAD (Pancreatic adenocarcinoma)

-   TCGA-ACC - Adrenocortical carcinoma)

-   TCGA-PCPG (Pheochromocytoma and Paraganglioma)

-   TCGA-KIRC (Kidney renal clear cell carcinoma)

-   TCGA-KICH (Kidney Chromophobe)

-   TCGA-KIRP (Kidney renal papillary cell carcinoma)

-   TCGA-COAD (Colon adenocarcinoma)

-   TCGA-READ (Rectum adenocarcinoma)

-   TCGA-OVCA(OV, Ovarian serous cystadenocarcinoma)

-   TCGA-UCEC (Uterine Corpus Endometrial Carcinoma)

-   TCGA-CESC (Cervical squamous cell carcinoma and endocervical
    adenocarcinoma)

-   TCGA-PRAD (Prostate adenocarcinoma)

-   TCGA-TGCT (Testicular Germ Cell Tumors)

-   TCGA-BLCA (Bladder Urothelial Carcinoma)

**Output of script:** One large csv data set containing all the
proteomic data of all patients of all cancer types. Additional clinical
patient data is included in case of further analysis. Proteins which
contain more than 50% of data values missing are removed.

### Downloading the data from the GDC website

Path to get to the data on the <https://portal.gdc.cancer.gov/> website:

-   Click on tab 'Projects'

-   Set the following filters - Program=TCGA, Data Category=proteome
    profiling

-   Click on the relevant TCGA-\_\_\_\_ project in the list

-   Click Explore Project Data

-   Click view files in repository

-   Set the following filters - Program=TCGA, Data Category=proteome
    profiling

-   Click 'Add all files to cart'

-   Go to cart

-   Download the files into the following folders

    -   Cart into the Biospecimen folder

    -   Clinical into the relevant Clinical folder

    -   Sample data into into the relevant sample_sheet folder

    -   Metadata into the relevant metadata folder

**NOTE:** When you download the cart and start putting them all in the
Biospecimen combined folder your system will ask if you would like to
replace the 'MANIFEST.txt' file just press 'skip this file' as it will
not be used in scripts.

**NOTE:** Download the Clinical file as a TSV not JSON.

## The script

### Organising the data into the appropriate folders

Making a list of the file paths of all the patients

```{r}


list_of_files <-
  list.files(
    path = "../Data/Raw_TCGA/Biospecimen",
    recursive = TRUE,
    pattern = "\\.tsv$",
    full.names = TRUE
  )

```

Obtaining the peptide names list

```{r}

peptide_list <- list()

for (i in list_of_files) {
  p <- read.table(file = i,
                  sep = '\t',
                  header = TRUE)
  for (n in 1:nrow(p)) {
    if (p[n, "peptide_target"] %in% peptide_list == FALSE) {
      peptide_list <- append(peptide_list, p$peptide_target)
    }
    else {
    }
  }
  
}

```

**NOTE:** To ensure that the protein list is the same for all patients
in this combined cancer type data set. This peptide list will be used as
the column headers.

Combining the clinical, and sample sheet data of the different cancer
types

```{r}
#Lung cancer - TCGA-LUAD study
setwd("../Data/Raw_TCGA/LUAD/Clinical")

clinical_data_LUAD <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)

exposure_data_LUAD <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)


```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/LUAD/Sample_sheet/")

sample_sheet_LUAD <-
  read.delim(file = "gdc_sample_sheet.2023-12-11.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
#Lung cancer - TCGA-LUSC study 
setwd("../Data/Raw_TCGA/LUSC/Clinical")

clinical_data_LUSC <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)

exposure_data_LUSC <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)


```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/LUSC/Sample_sheet/")

sample_sheet_LUSC <-
  read.delim(file = "gdc_sample_sheet.2024-01-17.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
#Liver cancer 
setwd("../Data/Raw_TCGA/LIHC/Clinical")

clinical_data_LIHC <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_LIHC <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/LIHC/Sample_sheet/")

sample_sheet_LIHC <-
  read.delim(file = "gdc_sample_sheet.2023-12-14.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
#Breast cancer - TCGA-BRCA 
setwd("../Data/Raw_TCGA/BRCA/Clinical")

clinical_data_BRCA <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_BRCA <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/BRCA/Sample_sheet/")

sample_sheet_BRCA <-
  read.delim(file = "gdc_sample_sheet.2024-01-01.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Colon adenocarcinoma - TCGA-COAD 
setwd("../Data/Raw_TCGA/COAD/Clinical")

clinical_data_COAD <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_COAD <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/COAD/Sample_sheet/")

sample_sheet_COAD <-
  read.delim(file = "gdc_sample_sheet.2024-01-19.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Rectum adenocarcinoma - TCGA-READ
setwd("../Data/Raw_TCGA/READ/Clinical")

clinical_data_READ <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_READ <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/READ/Sample_sheet/")

sample_sheet_READ <-
  read.delim(file = "gdc_sample_sheet.2024-01-19.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Kidney renal clear cell carcinoma - TCGA-KIRC 
setwd("../Data/Raw_TCGA/KIRC/Clinical")

clinical_data_KIRC <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_KIRC <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/KIRC/Sample_sheet/")

sample_sheet_KIRC <-
  read.delim(file = "gdc_sample_sheet.2024-01-19.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Ovarian serous cystadenocarcinoma - TCGA-OV (OVCA)
setwd("../Data/Raw_TCGA/OVCA/Clinical")

clinical_data_OVCA <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_OVCA <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/OVCA/Sample_sheet/")

sample_sheet_OVCA <-
  read.delim(file = "gdc_sample_sheet.2024-01-19.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Uterine Corpus Endometrial Carcinoma - TCGA-UCEC
setwd("../Data/Raw_TCGA/UCEC/Clinical")

clinical_data_UCEC <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_UCEC <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/UCEC/Sample_sheet/")

sample_sheet_UCEC <-
  read.delim(file = "gdc_sample_sheet.2024-01-19.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Head and Neck squamous cell carcinoma - TCGA-HNSC 
setwd("../Data/Raw_TCGA/HNSC/Clinical")

clinical_data_HNSC <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_HNSC <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/HNSC/Sample_sheet/")

sample_sheet_HNSC <-
  read.delim(file = "gdc_sample_sheet.2024-01-19.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Bladder Urothelial Carcinoma - TCGA-BLCA
setwd("../Data/Raw_TCGA/BLCA/Clinical")

clinical_data_BLCA <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_BLCA <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/BLCA/Sample_sheet/")

sample_sheet_BLCA <-
  read.delim(file = "gdc_sample_sheet.2024-01-19.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Glioblastoma multiforme - TCGA-GBM
setwd("../Data/Raw_TCGA/GBM/Clinical")

clinical_data_GBM <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_GBM <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/GBM/Sample_sheet/")

sample_sheet_GBM <-
  read.delim(file = "gdc_sample_sheet.2024-01-19.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Adrenocortical carcinoma - TCGA-ACC
setwd("../Data/Raw_TCGA/ACC/Clinical")

clinical_data_ACC <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_ACC <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/ACC/Sample_sheet/")

sample_sheet_ACC <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Brain Lower Grade Glioma - TCGA-LGG
setwd("../Data/Raw_TCGA/LGG/Clinical")

clinical_data_LGG <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_LGG <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/LGG/Sample_sheet/")

sample_sheet_LGG <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Cervical squamous cell carcinoma and endocervical adenocarcinoma - TCGA-CESC
setwd("../Data/Raw_TCGA/CESC/Clinical")

clinical_data_CESC <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_CESC <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/CESC/Sample_sheet/")

sample_sheet_CESC <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Esophageal carcinoma - TCGA-ESCA
setwd("../Data/Raw_TCGA/ESCA/Clinical")

clinical_data_ESCA <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_ESCA <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/ESCA/Sample_sheet/")

sample_sheet_ESCA <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Kidney Chromophobe - TCGA-KICH
setwd("../Data/Raw_TCGA/KICH/Clinical")

clinical_data_KICH <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_KICH <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/KICH/Sample_sheet/")

sample_sheet_KICH <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Kidney renal papillary cell carcinoma - TCGA-KIRP
setwd("../Data/Raw_TCGA/KIRP/Clinical")

clinical_data_KIRP <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_KIRP <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/KIRP/Sample_sheet/")

sample_sheet_KIRP <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Mesothelioma - TCGA-MESO
setwd("../Data/Raw_TCGA/MESO/Clinical")

clinical_data_MESO <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_MESO <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/MESO/Sample_sheet/")

sample_sheet_MESO <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Pancreatic adenocarcinoma - TCGA-PAAD
setwd("../Data/Raw_TCGA/PAAD/Clinical")

clinical_data_PAAD <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_PAAD <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/PAAD/Sample_sheet/")

sample_sheet_PAAD <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Pheochromocytoma and Paraganglioma - TCGA-PCPG
setwd("../Data/Raw_TCGA/PCPG/Clinical")

clinical_data_PCPG <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_PCPG <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/PCPG/Sample_sheet/")

sample_sheet_PCPG <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Prostate adenocarcinoma - TCGA-PRAD
setwd("../Data/Raw_TCGA/PRAD/Clinical")

clinical_data_PRAD <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_PRAD <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/PRAD/Sample_sheet/")

sample_sheet_PRAD <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Sarcoma - TCGA-SARC
setwd("../Data/Raw_TCGA/SARC/Clinical")

clinical_data_SARC <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_SARC <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/SARC/Sample_sheet/")

sample_sheet_SARC <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Skin Cutaneous Melanoma - TCGA-SKCM
setwd("../Data/Raw_TCGA/SKCM/Clinical")

clinical_data_SKCM <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_SKCM <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/SKCM/Sample_sheet/")

sample_sheet_SKCM <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Stomach adenocarcinoma - TCGA-STAD
setwd("../Data/Raw_TCGA/STAD/Clinical")

clinical_data_STAD <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_STAD <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/STAD/Sample_sheet/")

sample_sheet_STAD <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Testicular Germ Cell Tumors - TCGA-TGCT
setwd("../Data/Raw_TCGA/TGCT/Clinical")

clinical_data_TGCT <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_TGCT <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/TGCT/Sample_sheet/")

sample_sheet_TGCT <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Thymoma - TCGA-THYM
setwd("../Data/Raw_TCGA/THYM/Clinical")

clinical_data_THYM <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_THYM <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/THYM/Sample_sheet/")

sample_sheet_THYM <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

```{r}
# Thyroid carcinoma - TCGA-THCA
setwd("../Data/Raw_TCGA/THCA/Clinical")

clinical_data_THCA <-
  read.delim(file = "clinical.tsv", sep = '\t', header = TRUE)
exposure_data_THCA <-
  read.delim(file = "exposure.tsv", sep = '\t', header = TRUE)



```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

```{r}

setwd("../Data/Raw_TCGA/THCA/Sample_sheet/")

sample_sheet_THCA <-
  read.delim(file = "gdc_sample_sheet.2024-01-26.tsv", sep = '\t', header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

**NOTE:** For most studies follow_up, family_history, and pathology
datasets are empty, so I did not read them in.

Combining the data sets before exporting as a combined data set

```{r}
# Combining data sets
clinical_data <-
  rbind(clinical_data_LUSC,
        clinical_data_LUAD,
        clinical_data_LIHC,
        clinical_data_BRCA,
        clinical_data_COAD,
        clinical_data_READ,
        clinical_data_KIRC,
        clinical_data_OVCA,
        clinical_data_UCEC, 
        clinical_data_HNSC,
        clinical_data_BLCA,
        clinical_data_GBM,
        clinical_data_ACC,
        clinical_data_LGG,
        clinical_data_CESC,
        clinical_data_ESCA,
        clinical_data_KICH,
        clinical_data_KIRP,
        clinical_data_MESO,
        clinical_data_PAAD,
        clinical_data_PCPG,
        clinical_data_PRAD,
        clinical_data_SARC,
        clinical_data_SKCM,
        clinical_data_STAD,
        clinical_data_TGCT,
        clinical_data_THYM,
        clinical_data_THCA)
exposure_data <-
  rbind(exposure_data_LUSC,
        exposure_data_LUAD,
        exposure_data_LIHC,
        exposure_data_BRCA,
        exposure_data_COAD,
        exposure_data_READ,
        exposure_data_KIRC,
        exposure_data_OVCA,
        exposure_data_UCEC,
        exposure_data_HNSC,
        exposure_data_BLCA,
        exposure_data_GBM,
        exposure_data_ACC,
        exposure_data_LGG,
        exposure_data_CESC,
        exposure_data_ESCA,
        exposure_data_KICH,
        exposure_data_KIRP,
        exposure_data_MESO,
        exposure_data_PAAD,
        exposure_data_PCPG,
        exposure_data_PRAD,
        exposure_data_SARC,
        exposure_data_SKCM,
        exposure_data_STAD,
        exposure_data_TGCT,
        exposure_data_THYM,
        exposure_data_THCA)
sample_sheet <-
  rbind(sample_sheet_LUSC,
        sample_sheet_LUAD,
        sample_sheet_LIHC,
        sample_sheet_BRCA,
        sample_sheet_COAD,
        sample_sheet_READ,
        sample_sheet_KIRC,
        sample_sheet_OVCA,
        sample_sheet_UCEC,
        sample_sheet_HNSC,
        sample_sheet_BLCA,
        sample_sheet_GBM,
        sample_sheet_ACC,
        sample_sheet_LGG,
        sample_sheet_CESC,
        sample_sheet_ESCA,
        sample_sheet_KICH,
        sample_sheet_KIRP,
        sample_sheet_MESO,
        sample_sheet_PAAD,
        sample_sheet_PCPG,
        sample_sheet_PRAD,
        sample_sheet_SARC,
        sample_sheet_SKCM,
        sample_sheet_STAD,
        sample_sheet_TGCT,
        sample_sheet_THYM,
        sample_sheet_THCA)

```

Exporting the data frame as csv files into a global folder that will
contain the clinical, exposure, and sample sheet used for making the
combined cancer data set.

```{r}
write.csv(clinical_data,
          "../Data/Raw_TCGA/Clinical/clinical_data.csv",
          row.names = TRUE)
write.csv(exposure_data,
          "../Data/Raw_TCGA/Clinical/exposure_data.csv",
          row.names = TRUE)
write.csv(sample_sheet,
          "../Data/Raw_TCGA/Sample_sheet/sample_sheet.csv",
          row.names = TRUE)

```

### Creating a combined cancer dataset which contains protein expression and additional patient information which is of interest for the current study

Getting the clinical data

```{r}


setwd("../Data/Raw_TCGA/Clinical")

clinical_data <-
  read.csv(file = "clinical_data.csv", header = TRUE)
exposure_data <-
  read.csv(file = "exposure_data.csv", header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the clinical data.

Getting sample sheet

```{r}

setwd("../Data/Raw_TCGA/Sample_sheet")
sample_sheet <-
  read.csv(file = "sample_sheet.csv", header = TRUE)

```

**Warning explanation:** Changing the working directory of this chunk to
the folder which has the sample sheet data.

Making the global dataframe of patient identifiers (File.Names, File.ID,
case.submitter.id), and peptide expression

```{r}

# File name contains the File names of patients of all cancer types
File.Name <- basename(list_of_files)


#File.ID <- dirname(list_of_files)
#File.ID <- gsub("../Data/Raw_TCGA/Biospecimen/", "", File.ID)


# File name is found in the sample sheet data set, which also contains 'case.submitted.ID' (it is called the 'Case.ID' in the sample sheet) which found in the protein, clinical, and exposure data sets. 
case.submitter.id <- c()

for (i in File.Name) {
  if (i %in% sample_sheet$File.Name == TRUE) {
    r <-
      row.names(sample_sheet[which(sample_sheet == i, arr.ind = T)[, 1],])
    
    case.submitter.id <-
      append(case.submitter.id, sample_sheet$Case.ID[as.numeric(r)])
    
    
  }
  else{
    case.submitter.id <- append(case.submitter.id, "NA")
  }
}


#Obtaining the relevant clinical data 

stage <- c()

for (i in case.submitter.id) {
  if (i %in% clinical_data$case_submitter_id == TRUE) {
    r <-
      row.names(clinical_data[which(clinical_data == i, arr.ind = T)[, 1],])
    
    stage <- append(stage, suppressWarnings(clinical_data$ajcc_pathologic_stage[as.numeric(r[1])]))
    
  } else{
    stage <- append(stage, "NA")
  }
}


age <- c()

for (i in case.submitter.id) {
  if (i %in% clinical_data$case_submitter_id == TRUE) {
    r <-
      row.names(clinical_data[which(clinical_data == i, arr.ind = T)[, 1],])
    
    age <- append(age, suppressWarnings(as.numeric(clinical_data$age_at_index[as.numeric(r[1])])))
    
  } else{
    age <- append(age, "NA")
  }
}

days_to_death <- c()

for (i in case.submitter.id) {
  if (i %in% clinical_data$case_submitter_id == TRUE) {
    r <-
      row.names(clinical_data[which(clinical_data == i, arr.ind = T)[, 1],])
    
    days_to_death <- append(days_to_death, suppressWarnings(as.numeric(clinical_data$days_to_death[as.numeric(r[1])])))
    
  } else{
    days_to_death <- append(days_to_death, "NA")
  }
}

site_of_biopsy<- c()

for (i in case.submitter.id) {
  if (i %in% clinical_data$case_submitter_id == TRUE) {
    r <-
      row.names(clinical_data[which(clinical_data == i, arr.ind = T)[, 1],])
    
    site_of_biopsy <- append(site_of_biopsy, suppressWarnings(clinical_data$site_of_resection_or_biopsy[as.numeric(r[1])]))
    
  } else{
    site_of_biopsy <- append(site_of_biopsy, "NA")
  }
}

primary_diagnosis<- c()

for (i in case.submitter.id) {
  if (i %in% clinical_data$case_submitter_id == TRUE) {
    r <-
      row.names(clinical_data[which(clinical_data == i, arr.ind = T)[, 1],])
    
    primary_diagnosis <- append(primary_diagnosis, suppressWarnings(clinical_data$primary_diagnosis[as.numeric(r[1])]))
    
  } else{
    primary_diagnosis <- append(primary_diagnosis, "NA")
  }
}


project_id <- c()

for (i in case.submitter.id) {
  if (i %in% clinical_data$case_submitter_id == TRUE) {
    r <-
      row.names(clinical_data[which(clinical_data == i, arr.ind = T)[, 1],])
    
    project_id <- append(project_id, suppressWarnings(clinical_data$project_id[as.numeric(r[1])]))
    
  } else{
    project_id <- append(project_id, "NA")
  }
}


#Obtaining the relevant exposure data 

cigarettes_per_day <- c()

for (i in case.submitter.id) {
  if (i %in% exposure_data$case_submitter_id == TRUE) {
    r <-
      row.names(exposure_data[which(exposure_data == i, arr.ind = T)[, 1],])
    
    cigarettes_per_day <-
      append(cigarettes_per_day, suppressWarnings(as.numeric(exposure_data$cigarettes_per_day[as.numeric(r)])))
    
  } else{
    cigarettes_per_day <- append(cigarettes_per_day, "NA")
  }
}

years_smoked <- c()

for (i in case.submitter.id) {
  if (i %in% exposure_data$case_submitter_id == TRUE) {
    r <-
      row.names(exposure_data[which(exposure_data == i, arr.ind = T)[, 1],])
    
    years_smoked <-
      append(years_smoked, suppressWarnings(as.numeric(exposure_data$years_smoked[as.numeric(r)])))
    
  } else{
    years_smoked <- append(years_smoked, "NA")
  } 
}




# Making the dataframe
protein_data <-
  data.frame(
    case.submitter.id,
    project_id,
    age,
    primary_diagnosis,
    stage,
    site_of_biopsy,
    cigarettes_per_day,
    years_smoked,
    days_to_death,
    c(peptide_list)
  )

```

```{r}

# Creating a function which removes the full stop that was added to the end columns names to ensure that the protein names in this empty data frame I just created is the same as those found in the patients protein expression data sheets.


# Remove the first two characters from selected columns
cols_to_modify <- c(-1, -2, -3, -4, -5,-6,-7,-8, -9)
colnames(protein_data)[cols_to_modify] <- sub("^.{2}", "", colnames(protein_data)[cols_to_modify])

                                                                  


#removing the '.' at the end of the protein names
has_full_stop <- function(input_string) {
  # Use substr to extract the last character of the string
  last_character <-
    substr(input_string, nchar(input_string), nchar(input_string))
  
  # Check if the last character is a full stop
  return(last_character == ".")
}

library(stringr)
for (i in colnames(protein_data)) {
  if (has_full_stop(i) == TRUE) {
    #colnames(protein_data)<-str_sub(i,end = -2)
    colnames(protein_data)[which(colnames(protein_data) == i)] <-
      str_sub(colnames(protein_data)[which(colnames(protein_data) == i)], end = -2)
    
  }
  
}

rownames(protein_data) <- paste(File.Name)

```

**Warning suppression explanation:** when there is no value to pull out
there is a warning message "NAs introduced by coercion", this is want I
want to happen if there is no value to pull out so I suppressed this
warning. Then if there is anything other (e.g. '--') the else statement
puts a NA.

**NOTE:** At this point the data frame will contain the necessary number
of rows and columns. The additional patient information (e.g. age,
primary diagnosis etc.) columns will have values in them for each
patient. However, the protein expression data will not. Instead there is
just the protein name, this will be overwritten in the next code block
with the patients protein expression data.

**NOTE:** The next code block will take a longer time to process due
large amount of patients and due to the program having to look into each
of the patients 'Biospecimen' file to pull out all the necessary
information (it took me 1.45 hrs).

Filling in the protein expression data

```{r}

col_names_protein <- colnames(protein_data)
col_names_protein <- col_names_protein[c(-1, -2, -3, -4, -5,-6,-7,-8, -9)]

c <- c()
for (y in c(peptide_list)) {
  c <- c()
  y <- which(peptide_list == y)
  for (i in list_of_files) {
    n <- read.table(file = i,
                    sep = '\t',
                    header = TRUE)
    
    x <-
      row.names(n[which(n$peptide_target == peptide_list[y], arr.ind = T),])
    #the condition is saying if the data has the protein or not
    if (length(x) != 0) {
      if (is.na(n$protein_expression[as.numeric(x)]) == FALSE) {
        c <- append(c, n$protein_expression[as.numeric(x)])
      } else{
        c <- append(c, "NA")
      }
      
    } else{
      c <- append(c, "NA")
    }
  }
  y_index <-
    which(colnames(protein_data) == col_names_protein[y], arr.ind = T)
  protein_data[, y_index] <- c
  
  
}


```

### Removing proteins which have more that 50% of patients without data

Counting the number of NAs for each protein

```{r}

na_count <- c()
for (i in col_names_protein) {
  na_count <- append(na_count, (sum(protein_data[,i]=='NA') / 4581) * 100)
  
}

proteins_out <- c()
proteins_adjs <- c()

for (i in 1:length(na_count)) {
  if ((na_count[i] > 50) == TRUE) {
    proteins_out <- append(proteins_out, col_names_protein[i])
  }
  if ((na_count[i] < 50) == TRUE) {
    proteins_adjs <- append(proteins_adjs, col_names_protein[i])
    
  }
}

```

Removing the proteins from the data which have more than 50% of sample
values missing

```{r}
 
protein_data<-protein_data[,!(names(protein_data) %in% proteins_out)]


```

### Exporting the Full TCGA data which contains missing data values as a csv

Exporting the dataframe to a csv file

```{r}

write.csv(protein_data,
          "../Data/Processed_Data/Full_TCGA_data_missing_values.csv",
          row.names = TRUE)

```
